---
title: "Participants"
output: 
  html_document
---

<!-- Set general settings -->

```{r setup, include = FALSE}

# Set general settings for markdown file
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)


# Clear environment
rm(list = ls())


# Load packages
library(dplyr)      # for data manipulation
library(knitr)      # for integrating computing and reporting in markdown
library(kableExtra) # for customizing appearance of tables
library(foreign)    # for reading in SPSS file
library(psych)      # for describeBy function
library(tidyr)      # for gather and spread function


# Load functions
source("./functions/my_table_template.R")


# Turn off scientific notation
options(scipen = 999)
```
<br><br> 

## Sample Description
***

* The final sample consisted of 28 patients with OCD and 28 healthy control (HC) participants. 

* Patients with OCD and HC participants were individually matched for gender, age, and level of education.

* Patients were recruited from the specialized OCD outpatient clinic at Humboldt-Universität zu Berlin. All patients were diagnosed with OCD as the primary diagnosis, as assessed with the Structured Clinical Interview for DSM-IV (SCID-I; Wittchen et al., 1997).

* General inclusion criteria were age between 18 and 65 years and normal or corrected-to-normal vision. 

* Exclusion criteria for all participants were: lifetime diagnosis of any psychotic, bipolar, or substance-related disorder; use of benzodiazepines in the last week or of neuroleptic medication in the last three months; history of head trauma or neurological disease; any general exclusion criteria for tDCS. 

* Further exclusion criteria for healthy control participants were any current or past psychiatric disorder or current or past psychotherapeutic treatment. 

<br><br>

## Sample Characteristics
***

This table corresponds to Table 1 in the manuscript.

```{r demographics}

# Load data
participant_data <- read.spss("./data/Probandendaten.sav", to.data.frame = TRUE)
questionnaires   <- read.spss("./data/psychoEQExport.sav", to.data.frame = TRUE, add.undeclared.levels = "no")


# Select relevant variables
participant_data <- participant_data[, c("Code", "Alter", "Geschlecht",
  "Schuljahre", "YBOCS_gesamt", "YBOCS_Gedanken", "YBOCS_Handlungen")]
questionnaires   <- questionnaires[, c("CODE", "BD2SUMT0", "OCISUMT0")]


# Merge dataframes and rename columns
participant_data$Code <- gsub(" ", "", participant_data$Code, fixed = TRUE) # delete spaces for merging
demographics <- left_join(participant_data, questionnaires, by = c("Code" = "CODE")) %>%
  # Rename columns
  dplyr::rename("participant_id" = Code,
    "Age (years)"        = Alter,
    "Gender"             = Geschlecht,
    "Years of education" = Schuljahre,
    "Y-BOCS total score" = YBOCS_gesamt,
    "Y-BOCS obsessions"  = YBOCS_Gedanken,
    "Y-BOCS compulsions" = YBOCS_Handlungen,
    "BDI-II"             = BD2SUMT0,
    "OCI-R"              = OCISUMT0) %>%
  # Add group variable
  dplyr::mutate(group = as.factor(ifelse(substr(participant_id, 13, 13) == "C", "HC", "OCD"))) %>%
  # Exclude P_02 and C_02 and row with NA
  dplyr::filter(participant_id != "ModERN_tDCS_P02" & participant_id != "ModERN_tDCS_C02" & !is.na(Gender)) %>%
  # Select and reorder columns
  dplyr::select(c("group", "Age (years)", "Gender", "Years of education",  "BDI-II",  "OCI-R",
    "Y-BOCS total score",  "Y-BOCS obsessions",  "Y-BOCS compulsions"))



### Numeric variables

# Extract demographics per group
demographics_groups <- list()
demographics_groups <- psych::describeBy(demographics %>% dplyr::select(-c("Gender", "group")),
  demographics$group)
OCD <- as.data.frame(demographics_groups$`OCD`)
HC  <- as.data.frame(demographics_groups$`HC`)


# Create table with demographics formatted as "M (SD)"
demographics_groups <- as.data.frame(cbind(
  Characteristic = rownames(OCD),
  OCD = paste0(format(round(OCD$mean, 2), nsmall = 2), " (", format(round(OCD$sd, 2), nsmall = 2), ")"),
  HC  = paste0(format(round(HC$mean,  2), nsmall = 2), " (", format(round(HC$sd,  2), nsmall = 2), ")")
))


# Calculate t tests
t_tests <- demographics %>%
  dplyr::select("Age (years)", "Years of education",  "BDI-II",  "OCI-R", "group") %>%
  tidyr::gather(key = Characteristic, value = value, -group) %>%
  dplyr::group_by(group, Characteristic) %>%
  dplyr::summarise(value = list(value)) %>%
  tidyr::spread(group, value) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Characteristic) %>%
  dplyr::mutate("Test statistic" = t.test(unlist(HC), unlist(OCD))$statistic,
    df = t.test(unlist(HC), unlist(OCD))$parameter,
    p  = t.test(unlist(HC), unlist(OCD))$p.value) %>%
  dplyr::select(-c(HC, OCD))  %>%
  dplyr::ungroup()


# Combine demographics and t test results
demographics_numeric <- left_join(demographics_groups, t_tests, by = "Characteristic")



### Count variables (gender)

# Calculate chi-squared test and extract counts formatted as "female:male"
gender_test <- chisq.test(demographics$Gender, demographics$group)
gender      <- demographics %>% dplyr::group_by(group) %>% dplyr::count(Gender)
gender_OCD  <- paste0(gender[gender$group == "OCD" & gender$Gender == "weiblich", ]$n, ":",
                      gender[gender$group == "OCD" & gender$Gender == "männlich", ]$n)
gender_HC   <- paste0(gender[gender$group == "HC"  & gender$Gender == "weiblich", ]$n, ":",
                      gender[gender$group == "HC"  & gender$Gender == "männlich", ]$n)


# Combine demographics and chi-squared test
demographics_factorial <- as.data.frame(cbind(
  Characteristic = "Gender (female:male)",
  OCD = gender_OCD,
  HC  = gender_HC,
  "Test statistic" = gender_test$statistic,
  df = gender_test$parameter,
  p = gender_test$p.value
))



### Create and display table

# Re-order rows and format p values
table <- rbind(demographics_numeric[c(1), ],
  demographics_factorial,
  demographics_numeric[c(2:nrow(demographics_numeric)), ])
table[, c("df", "Test statistic", "p")] <- as.numeric(unlist(table[, c("df", "Test statistic", "p")]))
table[, c("p")] <- format.pval(table[, c("p")], eps = 0.001, digits = 3)


# Display table
my_table_template(table, caption = "Demographic and Clinical Characteristics of Patients With OCD and 
                  HC Participants ", digits = 2) %>%
  add_footnote("Note. Values are means with standard deviations in parentheses except for gender. 
               Years of education refer to primary and secondary education, not to higher education. 
               Test statistic refers to Welch’s t test except for gender, which was analyzed using
               a chi-squared test. BDI-II = Beck Depression Inventory-II; OCI-R = Obsessive-Compulsive 
               Inventory-Revised; Y-BOCS = Yale-Brown Obsessive Compulsive Scale.", notation = "none")
```
<br><br>

## References
***
Wittchen, H., Zaudig, M., & Fydrich, T. (1997). *Strukturiertes Klinisches Interview für DSM–IV. [User’s guide for the Structured Clinical Interview for DSM-IV–German modified version].* Hogrefe. 	
<br><br>

## Session Info
***
```{r session-info}

sessionInfo()
```
